### 2013-09-05 pystruct/paper.tex
added averaging, added SZLJSP algorithm

%TODO what about language / license?
\begin{table}[t]
\centering
\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccccc}
\toprule
Package       & Cutting Plane& SSGD & BCFW & Latent SSVM & Perceptron & ML\\
\cmidrule(r){1-1} \cmidrule(r){2-7}
\pystruct     & \x        & \x   & \x   & \x          & \x         & \o\\
\svmstruct    & \x        & \o   & \o   & \x          & \o         & \o\\
\sc{Dlib}     & \x        & \x   & \o   & \o          & \x         & \o\\
\sc{CRFsuite} & \o        & \o   & \o   & \o          & \x         & \x\\

\bottomrule
\end{tabularx}
    \caption{\label{table:comparision_algorithms}Comparison of learning
        algorithms implemented in popular structured prediction software
        packages. SSGD stochastic subgradient decent learning of SSVMs, BCFW for
    Block-Coordinate Frank-Wolfe, Latent SSVM for latent variable SSVMs, ML for
    maximum likelihood learning}
\end{table}

>>>>

%TODO what about language / license?
\begin{table}[t]
\centering
\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccccc}
\toprule
Package       & Cutting Plane& SSGD & STLJSP & Latent SSVM & Perceptron & ML\\
\cmidrule(r){1-1} \cmidrule(r){2-7}
\pystruct     & \x        & \x   & \x   & \x          & \x         & \o\\
\svmstruct    & \x        & \o   & \o   & \x          & \o         & \o\\
\sc{Dlib}     & \x        & \x   & \o   & \o          & \x         & \o\\
\sc{CRFsuite} & \o        & \o   & \o   & \o          & \x         & \x\\

\bottomrule
\end{tabularx}
    \caption{\label{table:comparision_algorithms}Comparison of learning
        algorithms implemented in popular structured prediction software
        packages. SSGD stochastic subgradient decent learning of SSVMs, STLJSP is
        as described in \secref{dual_coordinate_descent}, Latent SSVM for latent variable SSVMs, ML for
        maximum likelihood learning.}
\end{table}

### 2013-09-05 structured_prediction/paper.tex
added averaging, added SZLJSP algorithm

Arguably the most straight-foward way to approach \Eqref{learning_equation} is
using subgradient descent.  In light of the complexity of solving the
loss-augmented prediction problem in \Eqref{loss_augmentation}, is is natural
to work in a stochastic setting (see \citet{ratliff2007online}).
Given a model through $\phi$ and a set of parameters $\theta$, a subgradient
considering a single training example $(x\hoch{i}, y\hoch{i})$
can be computed simply by solving the loss-augmented prediction problem:
\begin{align}
    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\\
    &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y)
\end{align}
The most commonly used update has the simple form
\begin{equation}
    \theta \leftarrow (1 - \eta_t) \theta - \eta_t C \left [\phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i})\right]
\end{equation}
Where $\eta_t$ is a sequence of learning rates.
%Theoretical convergence guarantees of \ldots can be given whenever \ldots,
but in practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
Many practitioners adopted the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{pegasos}:
\begin{equation}
    \eta_t = \frac{C}{t},
\end{equation}
which has been found to work well in many settings.
Implementation of the algorithm is straight-forward, but unfortunately
detecting convergence is often tricky.
It is possible to use mini-batches instead of processing one sample at a time
to make use of multiple processors for inference. Unfortunately this negatively
affects the number of iterations needed, and did not provide a benefit in practice.

@@

\subsection{The STLJSP algorithm}\seclabel{dual_coordinate_descent}

Very recently, two groups independently derived a very performant new algorithm.
\citet{shalev2012proximal} investigated stochastic dual coordinate descent applied
to structured support vector machines. \citet{lacoste2012block} started from the Frank-Wolfe algorithm %TODO details
and derived a block-coordinate version, where each block corresponds to the
constraints associated with a single training example.  Both groups showed that
closed-form line search is possible, and, surprisingly ended up with the exact
same algorithm and closely related convergence guarantees.  The algorithm
processes a single training example at a time, and can be applied while
remaining completely in the primal domain, making it applicable to very large
datasets.  We will refer to the algorithm as STLJSP, after the authors of the
two papers, to honor the independent development of the two views of the
method.
%
The STDJSP algorithm has properties that are very simliar to
stochastic subgradient descent, but with two main advantages:
\begin{description}
\item[Stopping Criterion] Both views of STLJSP give rise to a dual objective,
    which can be used as a theoretically sound stopping criterion.
\item[Learning Rate] While there are several theoretical results on choosing
    the learning rate in stochastic subgradient descent, choosing a concrete
    schedule is often problematic in practice. By using analytic line-search,
    STLJSP removes the need for any learning rate parameter, making application
    of the algorithm in practice much simpler.
\end{description}

The STLJSP algorithm is usually used with weighted averaging, as this is known
to improve performance in similar stochastic algorithms, and empirically
improves convergence. % TODO weighted averaging here? or at SGD?

>>>>

Arguably the most straight-foward way to approach \Eqref{learning_equation} is
using subgradient descent.  In light of the complexity of solving the
loss-augmented prediction problem in \Eqref{loss_augmentation}, is is natural
to work in a stochastic setting (see \citet{ratliff2007online}).
Given a model through $\phi$ and a set of parameters $\theta$, a subgradient
considering a single training example $(x\hoch{i}, y\hoch{i})$
can be computed simply by solving the loss-augmented prediction problem:
\begin{align}
    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\\
    &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y)
\end{align}
The most commonly used update has the simple form
\begin{equation}
    \theta_{t+1} = (1 - \eta_t) \theta_t - \eta_t C \left [\phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i})\right]
\end{equation}
Where $\eta_t$ is a sequence of learning rates.
%Theoretical convergence guarantees of \ldots can be given whenever \ldots,
but in practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
Many practitioners adopted the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{shalev2011pegasos}:
\begin{equation}
    \eta_t = \frac{C}{t},
\end{equation}
which has been found to work well in many settings.
\citet{shalev2011pegasos} showed that this schedule achieves a convergence rate of $O(\frac{\ln T }{T})$.
\citet{lacoste2012block} and \citet{shamir2012stochastic} recently showed independently that
a rate of $O(\frac{1}{T})$ can be achieved using a novel averaging scheme:
\begin{equation}
    \bar{\theta}_{T} = \frac{2}{(T+1)(T+2)} \sum_{t=0}^T(t+1) \theta_t.
\end{equation}
This $t$-weighted averaging can be computed on-the-fly as
\begin{equation}\eqlabel{averaging}
    \bar{\theta}_{t} = \frac{t}{t + 2} \bar{\theta}_t + \frac{2}{t+2} \theta_{t+1}
\end{equation}

Implementation of the stochastic subgradient algorithm (with or without
averaging) is straight-forward, but unfortunately detecting convergence is
often tricky.
It is possible to use mini-batches instead of processing one sample at a time
to make use of multiple processors for inference. Unfortunately this negatively
affects the number of iterations needed, and did not provide a benefit in practice.

@@

%TODO BMRM?

\subsection{The SZLJSP algorithm}\seclabel{dual_coordinate_descent}
\begin{algorithm*}[t]
    \caption{Shalev-Shwartz Zhang Lacoste-Julien Jaggi Schmidt Pletscher\label{alg_szljsp}}
    \begin{algorithmic}[1]
        \Require training samples $\{ (x\hoch{i}, y\hoch{i}), \dots, (x\hoch{i}, y\hoch{i})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
        \Ensure parameters $\theta$
        \State $\theta_0, \theta_0\hoch{i}, \bar{\theta}_0  \leftarrow \mathbf{0}, \ell_0, \ell_0\hoch{i} \leftarrow 0 $
        \State $t \leftarrow 0$
        \Repeat
            \State $t \leftarrow t + 1$
            \State Pick $i$ uniformly at random from \{1, \dotsc, k\}.
            \State $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] $
            \State Compute parameter and loss updates based on sample $i$:\\
                $\theta_s \leftarrow \frac{C}{n} \phi(\hat{y})$
                $l_s \leftarrow \frac{C}{n} \delta(y\hoch{i}, \hat{y})$
            \State Compute optimum step size $\eta$:\\
                $\eta \leftarrow \frac{(\theta_t\hoch{i} - \theta_s)^T \theta_t + C (\ell_s - \ell_k\hoch{i} )}{\lVert \theta_t\hoch{i} - \theta_s \rVert^2}$ and clip to $[0, 1]$
            \State Update per-sample parameters and loss estimate:\\
                $\theta_{t+1}\hoch{i} \leftarrow (1 - \eta) \theta_{t+1}\hoch{i} + \eta \theta_s$\\
                $\ell_{t+1}\hoch{i} \leftarrow (1 - \eta) \theta_{t+1}\hoch{i} + \eta \ell_s$
            \State Update global parameters and loss estimate:\\
                $\theta_{t+1} \leftarrow \theta_{t+1} + \theta_{t}\hoch{i} - \theta_{t+1}\hoch{i}$\\
                $\ell_{t+1} \leftarrow \ell_{t+1} + \ell_{t}\hoch{i} - \ell_{t+1}\hoch{i}$
            \State Compute the weighted running average:\\
            $\bar{\theta}_{t+1} = \frac{k}{k + 2} \bar{\theta}_k + \frac{2}{k + 2}\theta_{k+1}$
            \vspace{-5mm}
        \Until
    \end{algorithmic}
\end{algorithm*}

Very recently, two groups independently derived a very performant new algorithm.
\citet{shalev2012proximal} investigated stochastic dual coordinate descent applied
to structured support vector machines. \citet{lacoste2012block} started from the Frank-Wolfe algorithm %TODO details
and derived a block-coordinate version, where each block corresponds to the
constraints associated with a single training example.  Both groups showed that
closed-form line search is possible, and, surprisingly ended up with the exact
same algorithm and closely related convergence guarantees.  The algorithm
processes a single training example at a time, and can be applied while
remaining completely in the primal domain, making it applicable to very large
datasets.  We will refer to the algorithm as SZLJSP, after the authors of the
two papers, to honor the independent development of the two views of the
method.
%
The SZLJSP algorithm has properties that are very simliar to
stochastic subgradient descent, but with two main advantages:
\begin{description}
\item[Stopping Criterion] Both views of SZLJSP give rise to a dual objective,
    which can be used as a theoretically sound stopping criterion.
\item[Learning Rate] While there are several theoretical results on choosing
    the learning rate in stochastic subgradient descent, choosing a concrete
    schedule is often problematic in practice. By using analytic line-search,
    SZLJSP removes the need for any learning rate parameter, making application
    of the algorithm in practice much simpler.
\end{description}

In practice, the SZLJSP algorithm is usually used with the weighted averaging
described in \Eqref{averaging}, as this is known to improve performance in the
closely related stochastic subgradient algorithms, and empirically improves
convergence.

The detailed procedure as given by \citet{lacoste2012block} is shown in Algorithm~\ref{alg_szljsp}.
